Current and future constraints on primordial magnetic fields

### 2016-10-20 pmf.tex
a

\newcommand{\sqdeg}{deg$^2$ }
\newcommand{\omb}{\ensuremath{\Omega_b h^2}}
\newcommand{\omc}{\ensuremath{\Omega_c h^2}}
\newcommand{\clpp}{\ensuremath{C_{L}^{\phi\phi}}}
\newcommand{\cpmf}{\ensuremath{C_{\ell}^{\rm pmf}}}
\newcommand{\apmf}{\ensuremath{A_{\rm pmf}}}
\newcommand{\bpmf}{\ensuremath{B_{\rm 1\,Mpc}}}
\newcommand{\alens}{\ensuremath{A_{\rm lens}}}
\newcommand{\lcdm}{\ensuremath{\Lambda}CDM}
\newcommand{\fermilat}{\textit{Fermi}-LAT}

@@

\subsection{Fiducial survey parameters}

@@

@@

>>>>

\newcommand{\sqdeg}{deg$^2$ }
\newcommand{\omb}{\ensuremath{\Omega_b h^2}}
\newcommand{\omc}{\ensuremath{\Omega_c h^2}}
\newcommand{\clpp}{\ensuremath{C_{L}^{\phi\phi}}}
\newcommand{\cpmf}{\ensuremath{C_{\ell}^{\rm pmf}}}
\newcommand{\apmf}{\ensuremath{A_{\rm pmf}}}
\newcommand{\bpmf}{\ensuremath{B_{\rm 1\,Mpc}}}
\newcommand{\alens}{\ensuremath{A_{\rm lens}}}
\newcommand{\lcdm}{\ensuremath{\Lambda}CDM}
\newcommand{\nrun}{\ensuremath{n_{\rm run}}}
\newcommand{\neff}{\ensuremath{N_{\rm eff}}}
\newcommand{\mnu}{\ensuremath{\sum m|\nu}}
\newcommand{\fermilat}{\textit{Fermi}-LAT}

@@

\subsection{Methods and Experimental Parameters}

We use Fisher matrices to forecast the constraints on PMFs possible from each generation of experiment. 
Fisher matrices make it easy to combine different experiments and forecast the final parameter constraints \citep{fisher}. 
According to the Cramer-Rao bound, the Fisher information is the lower bound on the uncertainty

We include two external datasets in all forecasts. 
The first are the expected measurements of the  TT, TE and EE spectra from the \planck{} satellite. 
We include \planck{} TT information in the multipole range $2\le \ell \le 3000$. 
Due to the importance of galactic foreground removal at large scales in polarization, we only use TE and EE information starting from $\ell = 30$. 
A prior on the optical depth of 0.005 is added to account for the expected optical depth constraint from the missing multipoles. 
Second, we include a 1\% external measurement of the Hubble constant, such as would be expected from the Taipan experiment \citep{taipan}.

One concern with combining Fisher matrices from different experiments is double-counting modes due to overlapping sky coverage. 
We sidestep this issue for the TT or TE spectra, by only using the \planck{} measurements. 
We do not expect the future experiments to substantially improve upon \planck{} in these spectra. 
We throw away \planck{} EE data in the overlap region by appropriately increasing the forecast \planck{} EE uncertainties. 
The exact increase depends on the assumed stage III or IV survey area.
We do not include \planck{} BB information in the forecasts. 
We ignore overlaps between the stage III or IV experiments, with the note that the overlap between a Chilean experiment like the Simons Array and South Pole experiment like SPT-3G should be small.

\begin{table*}[h]
\begin{center}
\caption{\label{tab:experiments} Assumed survey parameters}
\small
\begin{tabular}{l || c c c c c }
Experiment & Sky coverage & Polarized Noise level  & 1/f knee & Beam FWHM \\
& &($\mu$K-arcmin)&&(arcmin.)\\
\hline
Stage III: & & & & \\

~~~~SPT-3G & 6\% & 3.0 & 200 & 1.2 \\
~~~~Simons Array & 36\% & 9.5 & 200 & 3.5 \\ 
\\
%\hline
CMB Stage IV & 55\% & 1.3 & 100 & 4.0 \\
\end{tabular}
\tablecomments{ 
Key numbers about the planned stage III and IV experiments. 
The sky coverage percentages are after galactic cuts. 
Unless otherwise noted,  the Fisher matrix forecasts in this work use these numbers. 
All forecasts also include beam and calibration uncertainty as noted in the text. 
} \normalsize
\end{center}
\end{table*}

We list the assumed survey areas, noise levels and beam sizes for each experiment in Table~\ref{tab:experiments}. 
We also assume a 5\% uncertainty on the beam FWHM and a 1\% power calibration uncertainty; we have tested relaxing or tightening the beam and calibration uncertainty and find these calibration terms have a negligible impact on the PMF constraints.

We consider the constraints on PMFs in two cosmologies. 
In all cases, we also marginalize over unknown Poisson EE and BB  terms due to polarized extragalactic sources. 
Generally, we report constraints on the PMFs as one of several extensions to LCDM:  \lcdm{}  +r + \nrun{} + \neff{} + \mnu{}+ \apmf. 
However, in order to illustrate the degree to which parameter degeneracies are limiting the inferred constraints,  we also show constraints from an aggressive model in which the PMF power is the only extension to \lcdm{}:  \lcdm{}  + \apmf.

@@

The experiments that will begin taking data next year will dramatically improve constraints on the PMF power. 
With the minimal cosmological model, the 1-sigma forecasts for \planck{}+\ho{} is $\sigma(\apmf)=0.47$, slightly worse than indicated by the upper limits found with the full \planck{} likelihood. 
The uncertainty falls nearly 400-fold by adding the two stage III experiments in this simplified model. 
Including all eleven parameters marginally weakens the \planck+\ho{} constraints on \apmf{} by 23\%. 
Parameter degeneracies degrade the Stage III constraint by a substantial factor (x2.5). 
Still in this 11-parameter model, the addition of stage III CMB experiments improves the \apmf{} uncertainty by a factor of $\sim$180 to  $\sigma(\apmf)=0.0031$. 
We can expect substantially tighter constraints on PMFs by the end of the decade.

@@

The primary motivation behind the proposed CMB stage IV experiment is to search for inflationary gravitational waves and neutrino masses, however, the experiment should also enable extremely sensitive searches for PMFs.

 extremely powerful
